Patchwork: Incremental Computation for Static Analysis

### 2018-09-24 paper.tex
new paper.tex doc

%% Title information
\title[Short Title]{Meta Analysis for Optimal Run-time Checking}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

Static analysis has several common failure modes where it is unable to discharge important safety and correctness assertions about programs. Common behaviors such as complicated loops, randomness/(stochastic behavior), and dynamic data loads (external interaction) frequently stymie static analyses. On the other hand, verifying these assertions via runtime checks requires a programmer to correctly reason about how and when to perform these checks, which can be difficult. We propose a middle-ground approach which performs \textit{online} static analyses to verify programs, and propose a tool which determines the optimal location to discharge assertions as early as possible.

@@

Interval analysis (w/common static failure modes)
Anderson's analysis ("")
Typechecking ("", ref. GPCE)

\section{Approach: meta-analysis}

Req: program, programmer-provided annotation (to be discharged), analysis to be used to discharge the annotation, semantics for stepping program
Tool:
    Run analysis at each step, identify stoplight for annotation (red/yellow/green)
        Modify analysis to handle running in-program? (several options)
    Identify steps where stoplight result \textit{changes}
    Return to user: location of optimal check, or perhaps "can't verify"

@@

go back through motivating examples

@@

\textit{Incremental computation} fixes the efficiency problem, move the model from pre-production checking/program modification to in-production online verification

@@

>>>>

%% Title information
\title[Short Title]{Patchwork: Incremental Computation for Static Analysis}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

Programmers who use static analyses in their normal workflow will frequently demand the results of these analyses multiple times. For example, when working on code in an IDE, a programmer may make multiple small edits and request the results of an anlysis after each one. Since these edits are likely to be minor, performing the entire analysis from scratch each time is inefficient. This is a prime application for \textit{incremental computation} (insert typical IC story).

Incremental computation is reasonably well understood for standard Datalog, which can express a small class of analyses. However, more expressiveness is needed to implement complex lattice-based analyses, particularly on infinite-height lattices, such as an interval analysis. We present the Patchwork framework, which extends the Datalog-based approach to allow for incremental implementations of general static analyses.

@@

Parity analysis (requires lattice)\\
Interval analysis (requires widening)

\section{Approach: fungilang + lattice?}

Present fungilang semantics

@@

go back through motivating examples: parity analysis

\begin{lstlisting}
1 x = 1;
2 while x < rand()
3   x++;
\end{lstlisting}

\section{Evaluation}

We evaluated our approach on a spread of analyses and target programs.

Analyses:\\
Simple interval? Points-to? Interprocedural?

Programs:\\
IncA benchmarks? Flix benchmarks? Both?

Edit model:\\
Adopt IncA approach? Commit history? Other?

We compare to {Flix/IncA}?

@@

Interprocedural? Or doing that now?

\section{Related Work}

Incremental Datalog framework (which?):

Flix: the Patchwork framework extends the work of the Flix authors. Specifically, Flix provides a Datalog extension with the ability to reason about lattices. Patchwork extends work done on incremental Datalog evaluation, to, effectively, incremental Flix evaluation. Notably, Flix evaluates (large numbers) of rules in the course of (example). This includes significant redundancy, due to (facts about their dependency analysis). Our approach automatically tracks these dependencies and improves performance by (numbers).

IncA: the IncA papers gives an algorithm for performing incremental evaluation of lattice-aware Datalog-style programs. The approach takes advantage of properties of monotonic operators on lattices to quickly update aggregations and Datalog relations. However, the monotonicity restriction rules out important program analyses, such as (those that depend on widening). The IncA framework cannot express our second worked example. Additionally, (performance).

@@

Existing state-of-the-art approaches to incremental program analysis are not fully general. Incremental Datalog engines are restricted to the simple declarative model, and even more complex approaches such as IncA restrict the types of operators an analysis user can work with. The Patchwork framework provides a fully general (is this true?) approach to incremental static analysis. Our use of Adapton also improves efficiency by (numbers). This represents a significant advance in the state of incremental static analysis.
